Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-24 src/tex/ms.tex
did we have to shorten the abstract? remove final sentence.
semi-analytical analysis: lower limit 10^-15

\begin{abstract}
A wide variety of scenarios for the origin of life have been proposed, with many influencing the prevalence and distribution of biosignatures across exoplanet populations.
This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
Using Bayesian modeling and the \bioverse\ survey simulator, we constrain the probability of obtaining strong evidence for or against this ``UV Threshold Hypothesis'' with future biosignature surveys.
Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
Surveys targeting a wide range of fluxes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
Our findings demonstrate the potential of exoplanet surveys to test origins-of-life hypotheses.
 Beyond specific scenarios, this work underscores the broader value of realistic survey simulations for future observatories (e.g., HWO, LIFE, ELTs, Nautilus) in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
% This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
\end{abstract}

@@

Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution (but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors).}
Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
Under Hypothesis $H_1$, there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that

@@

\subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution}.
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence}
        \caption{Obtaining true strong evidence with different sample sizes. Left: Probability to reach given evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). Vertical lines denote thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of obtaining true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: \rev{for $n = \var{semian_Nsamp2}$, \SI{14}{\percent} (\SI{16}{\percent}) of} random samples permit conclusive inference (strong true evidence) \rev{under $H_\mathrm{1}$ ($H_\mathrm{null}$).}

@@

\section{\rev{Alternative priors for the abiogenesis rate}}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}

>>>>

\begin{abstract}
A wide variety of scenarios for the origin of life have been proposed, with many influencing the prevalence and distribution of biosignatures across exoplanet populations.
This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
Using Bayesian modeling and the \bioverse\ survey simulator, we constrain the probability of obtaining strong evidence for or against this ``UV Threshold Hypothesis'' with future biosignature surveys.
Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
Surveys targeting a wide range of fluxes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
Our findings demonstrate the potential of exoplanet surveys to test origins-of-life hypotheses.
 Beyond specific scenarios, this work underscores the broader value of realistic survey simulations for future observatories (e.g., HWO, LIFE, ELTs, Nautilus) in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
\end{abstract}

@@

Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution} \revv{over the range $[10^{-15}, 1]$~\citep[][but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors]{Balbi2023a}}.
Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
Under Hypothesis $H_1$, there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that

@@

\subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution \revv{over the range $[10^{-15}, 1]$}}.
For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
In the majority of cases, the outcome of the survey will be inconclusive.
\begin{figure*}
    \script{semian_true_evidence.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/semian_true_evidence}
        \caption{Obtaining true strong evidence with different sample sizes. Left: Probability to reach given evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed). Vertical lines denote thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100. Right: Probability of obtaining true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
        \label{fig:semian_true_evidence}
    \end{centering}
\end{figure*}
The situation improves with larger samples: \rev{for $n = \var{semian_Nsamp2}$, \SI{14}{\percent} (\SI{16}{\percent}) of} random samples permit conclusive inference (strong true evidence) \rev{under $H_\mathrm{1}$ ($H_\mathrm{null}$).}

@@

\section{\rev{Alternative priors for the abiogenesis rate}}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior \revv{over the range $[10^{-15}, 1]$} for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}
